\documentclass[11pt, oneside]{article}
\usepackage[letterpaper, margin=2cm]{geometry}
\usepackage{MATH667}
\usepackage{booktabs}

\begin{document}
\noindent \textbf{\Large{Caleb Logemann \\
MATH667 Hyperbolic Partial Differential Equations \\
Homework 7
}}

%\lstinputlisting[language=MATLAB]{H01_23.m}
\begin{enumerate}
  \item % #1
    Derive the following 3\textsuperscript{rd} order accuracy reconstruction
    for finite volume methods.
    \begin{align*}
      u_{j+1/2}^- &= -\frac{1}{6} \bar{u}_{j-1} + \frac{5}{6}\bar{u}_j + \frac{1}{3}\bar{u}_{j+1} \\
      u_{j+1/2}^+ &= \frac{1}{3} \bar{u}_{j} + \frac{5}{6}\bar{u}_{j+1} - \frac{1}{6}\bar{u}_{j+2}
    \end{align*}

    For this problem we would like to construct a quadractic polynomial $p(x)$
    such that $p$ matches the cell average on intervals $I_{j-1}$, $I_j$, and
    $I_{j+1}$.
    If we set
    $p(x) = a\p{x - x_{j+1/2}}^2 + b\p{x - x_{j-1/2}} + c$, then
    $p(x_{j-1/2}) = c$.
    We can then solve the following three equations for $c$.
    \begin{align*}
      \frac{1}{h} \dintt{x_{j-3/2}}{x_{j-1/2}}{p(x)}{x} &= \bar{u}_{j-1} \\
      \frac{1}{h} \dintt{x_{j-1/2}}{x_{j+1/2}}{p(x)}{x} &= \bar{u}_{j} \\
      \frac{1}{h} \dintt{x_{j+1/2}}{x_{j+3/2}}{p(x)}{x} &= \bar{u}_{j+1}
    \end{align*}
    These equation can be simplified as follows.
    \begin{align*}
      \frac{1}{3h}a \p{-h^3 + 8h^3} + \frac{1}{2h}b\p{h^2 - 4h^2} + c &= \bar{u}_{j-1} \\
      \frac{1}{3h}a \p{h^3} + \frac{1}{2h}b\p{-h^2} + c &= \bar{u}_{j} \\
      \frac{1}{3h}a \p{h^3} + \frac{1}{2h}b\p{h^2} + c &= \bar{u}_{j+1}
    \end{align*}
    Simplifying gives
    \begin{align*}
      \frac{7h^2}{3}a - \frac{3h}{2}b + c &= \bar{u}_{j-1} \\
      \frac{h^2}{3}a - \frac{h}{2}b + c &= \bar{u}_{j} \\
      \frac{h^2}{3}a + \frac{h}{2}b + c &= \bar{u}_{j+1}.
    \end{align*}
    Subtracting three times the second equation to the first equation and adding the last two equations gives
    \begin{align*}
      \frac{4h^2}{3}a - 2c &= \bar{u}_{j-1} - 3\bar{u}_j \\
      \frac{2h^2}{3}a + 2c &= \bar{u}_j + \bar{u}_{j+1}.
    \end{align*}
    Subtracting 2 times the second equation from the first gives
    \begin{align*}
      -6c &= \bar{u}_{j-1} - 5\bar{u}_j - 2\bar{u}_{j+1} \\
      c &= -\frac{1}{6}\bar{u}_{j-1} + \frac{5}{6}\bar{u}_j + \frac{1}{3}\bar{u}_{j+1}
    \end{align*}
    Thus we have for the first reconstruction that
    \[
      u^-_{j+1/2} = p(x_{j+1/2}) = c = -\frac{1}{6}\bar{u}_{j-1} + \frac{5}{6}\bar{u}_j + \frac{1}{3}\bar{u}_{j+1}
    \]

    We can now do this process again on the intervals $I_j$, $I_{j+1}$, and $I_{j+2}$.
    This only changes one of the equations.
    We now have
    \begin{align*}
      \frac{1}{h} \dintt{x_{j+3/2}}{x_{j+5/2}}{p(x)}{x} &= \bar{u}_{j+2} \\
      \frac{1}{3h}a \p{8h^3 - h^3} + \frac{1}{2h}b\p{4h^2 - h^2} + c &= \bar{u}_{j+2} \\
      \frac{7h^2}{3}a + \frac{3h}{2}b + c &= \bar{u}_{j+2}
    \end{align*}
    We must now solve the following three equations for $c$
    \begin{align*}
      \frac{h^2}{3}a - \frac{h}{2}b + c &= \bar{u}_{j} \\
      \frac{h^2}{3}a + \frac{h}{2}b + c &= \bar{u}_{j+1} \\
      \frac{7h^2}{3}a + \frac{3h}{2}b + c &= \bar{u}_{j+2}.
    \end{align*}
    Adding the first two equations and subtracting 3 times the second
    equation from the last equation gives
    \begin{align*}
      \frac{2h^2}{3}a + 2c &= \bar{u}_{j} + \bar{u}_{j+1} \\
      \frac{4h^2}{3}a + -2c &= \bar{u}_{j+2} - 3\bar{u}_{j+1}.
    \end{align*}
    Subtracting 2 times the first equation from the second equation gives
    \begin{align*}
      -6c &= -2\bar{u}_{j} - 5\bar{u}_{j+1} + \bar{u}_{j+2} \\
      c &= \frac{1}{3}\bar{u}_{j} + \frac{5}{6}\bar{u}_{j+1} - \frac{1}{6}\bar{u}_{j+2}
    \end{align*}
    This is the second reconstruction
    \[
      u^+_{j+1/2} = p(x_{j+1/2}) = c = \frac{1}{3}\bar{u}_{j} + \frac{5}{6}\bar{u}_{j+1} - \frac{1}{6}\bar{u}_{j+2} \\
    \]

  \item % #2
    Consider to solve the 1D scalar conservation law.
    Show the 3rd order finite volume MUSCL scheme is TVD.
    Use Harten's Thereom.

    First I will write out explicitly what the 3rd order MUSCL scheme is.

  \item % #3
    Solve two-phase flow nonconver Buckley-Leverett equation with
    3\textsuperscript{rd} order finite volume MUSCL scheme.
    \[
      u_t + \p{\frac{2u^2}{2u^2 + (1 - u)^2}}_x = 0
    \]
    with initial conditions
    \[
      u(x, 0) =
      \begin{cases}
        1 & x \le 0 \\
        0 & x > 0
      \end{cases}
    \]
    The solution consists of a rarefaction wave connecting with a shock, ref
    to Figure 4.7 on page 49.
    Simulate the evolution of the solution to $T = 1.0$ with total mesh
    $N = 100$.
    Output the simulate with one symbol per cell on the figure.
    Use exact solutions at the boundary (ghost cells) as the given boundary
    conditions.

\end{enumerate}
\end{document}
